---
title: "R Notebook"
author: Aman Agarwal - A20414691
output: 
  html_notebook:
    toc: true
    toc_depth: 1
---


# Including necessary libraries
```{r}
library(textreuse)
library(stringr)
library(lsa)
library(naniar)
library(dplyr)
library(DataCombine)
library(tm)

```

# Problem 2.1
```{r}
movies <- read.csv("movies.csv", header = TRUE, sep = ',')
ratings <- read.csv("ratings.csv", header = TRUE, sep = ',')
unique_userId <- unique(ratings['userId'])

for (userId in unique_userId$userId) {
  movieIds_for_particular_user <- ratings[ratings$userId == userId, 'movieId']
  movies_for_particular_user <- data.frame(movies[is.element(movies$movieId, movieIds_for_particular_user), 'title'])
  write.table(movies_for_particular_user, row.names = FALSE, col.names = FALSE, quote = FALSE, paste('hw4.movies/user',userId,'.txt', sep = ''))
}

files <- list.files("hw4.movies", full.names = TRUE)
corpus <- TextReuseCorpus(files, tokenizer = tokenize_ngrams, n = 5, keep_tokens = TRUE)

```

# Problem 2.1 (a)
```{r}
total_number_of_documents <- length(corpus)
unique_corpus <- c()

for (i in 1:total_number_of_documents) {
  corpus_for_this_document <- corpus[[i]]$tokens
  unique_corpus <- union(unique_corpus, corpus_for_this_document)
}

cat("Number of rows = ", length(unique_corpus), " and columns = ", total_number_of_documents, sep = "")

```


# Problem 2.1 (b) (i)
```{r}
user20_corpus <- corpus[["user20"]]
cat("User20 has rated", str_count(user20_corpus$content, "\n") + 1, "movies")

```

# Problem 2.1 (b) (ii)
```{r}
cat("The first 5 shingles (tokens) for user20 are:\n")
print(user20_corpus$tokens[1:5])

```

# Problem 2.1 (c)
```{r}
pairwise_candidates_res <- pairwise_candidates(pairwise_compare(corpus, jaccard_similarity))

```

# Problem 2.1 (c) (i)
```{r}
cat("Number of pairs with similarity score of atleast 0.60 is", nrow(pairwise_candidates_res[pairwise_candidates_res$score>=0.60, ]), "\n")

```

# Problem 2.1 (c) (ii)
```{r}
cat("Number of pairs with similarity score of atleast 0.50 is", nrow(pairwise_candidates_res[pairwise_candidates_res$score>=0.50, ]), "\n")

```

# Problem 2.1 (c) (iii)
```{r}
pairs_based_on_similarity_score <- pairwise_candidates_res[pairwise_candidates_res$score>=0.40, ]
cat("Number of pairs with similarity score of atleast 0.40 is", nrow(pairs_based_on_similarity_score), "\n")

```

# Problem 2.1 (c) (iv)
```{r}
print(pairs_based_on_similarity_score)

```

# Problem 2.1 (d) (i)
```{r}
h <- 5
b <- 5
prob <- lsh_probability(h = h, b = b, s = 0.60)
cat("Minimum number of minhash signatures =", h, "and number of LSH bands =", b)

```

# Problem 2.1 (d) (ii)
```{r}
minhash <- minhash_generator(n = h, seed = 100)
minhash_corpus <- TextReuseCorpus(files, tokenizer = tokenize_ngrams, n = 5, keep_tokens = TRUE, minhash_func = minhash)
user20_minhash_corpus <- minhash_corpus[["user20"]]
cat("For user ID 20, the first five minhashes are:\n")
print(user20_minhash_corpus$minhashes[1:5])

```

# Problem 2.1 (e)
```{r}
buckets <- lsh(minhash_corpus, bands = b)
candidates <- lsh_candidates(buckets)
res <- lsh_compare(candidates, minhash_corpus, jaccard_similarity)

```

# Problem 2.1 (e) (i)
## We do not expect any candidate pairs that have a Jaccard similarity between 0.60 and 0.50 because in 2.1(c)(i) and (ii) part we have already seen using brute force method, that number of candidate pairs with jaccard similarity between 0.5 and 0.6 is zero. Same result was expected out of LSH also.
```{r}
cat("Number of pairs with similarity score between 0.60 and 0.50 is", nrow(res[res$score<=0.60 & res$score>=0.50, ]), "\n")

```

# Problem 2.1 (e) (ii)
```{r}
pairs_lsh <- res[res$score>=0.40, ]
print(pairs_lsh)

```

# Problem 2.1 (e) (iii)
## Yes the pairs and there scores are same as in 2.1(c)(iv)
```{r}

```

# Problem 2.1 (e) (iv)
```{r}
lsh_comparisons <- nrow(res)
brute_force_comparisons <- nrow(pairwise_candidates_res)
cat("Work saved = ", ((brute_force_comparisons - lsh_comparisons)/brute_force_comparisons)*100,"%", sep = "")

```

# Problem 2.2
```{r}
A_number = 20414691
user_profile_id = A_number %% 671

# Getting the movieIds watched by the user 187 and initializing the genres list
movieIds_for_user187 <- ratings[ratings$userId == user_profile_id, 'movieId']
genres <- c('Action', 'Adventure', 'Animation', 'Children', 'Comedy', 'Crime', 'Documentary', 'Drama', 'Fantasy', 'Film-Noir', 'Horror', 'IMAX', 'Musical', 'Mystery', 'Romance', 'Sci-Fi', 'Thriller', 'War', 'Western', 'no genres listed')

random_samples_to_choose <- 10

# Creating User profile matrix
user_profile_matrix <- data.frame(matrix(nrow = length(movieIds_for_user187), ncol = length(genres)))
rownames(user_profile_matrix) <- movieIds_for_user187
colnames(user_profile_matrix) <- genres

for (movieId in movieIds_for_user187) {
  genres_list_of_particular_movie <- as.character(movies[movies$movieId == movieId, 'genres'])
  genres_split_of_particular_movie <- unlist(strsplit(as.character(genres_list_of_particular_movie), '|', fixed = TRUE))
  
  for(genre in genres_split_of_particular_movie) {
    user_profile_matrix[as.character(movieId), as.character(genre)] = 1
  }
}

# Creating the User profile vector from User profile matrix
user_profile_matrix[is.na(user_profile_matrix)] <- 0
user_profile_vector <- colMeans(user_profile_matrix)

# Getting the random movieIds
random_movieIds <- sample(movies[, 'movieId'], size = random_samples_to_choose, replace = FALSE)

# Creating Movie profile matrix
movie_profile_matrix <- data.frame(matrix(nrow = random_samples_to_choose, ncol = length(genres)))
rownames(movie_profile_matrix) <- random_movieIds
colnames(movie_profile_matrix) <- genres

for (movieId in random_movieIds) {
  genres_list_of_particular_movie <- as.character(movies[movies$movieId == movieId, 'genres'])
  genres_split_of_particular_movie <- unlist(strsplit(as.character(genres_list_of_particular_movie), '|', fixed = TRUE))
  
  for(genre in genres_split_of_particular_movie) {
    movie_profile_matrix[as.character(movieId), as.character(genre)] = 1
  }
}

movie_profile_matrix[is.na(movie_profile_matrix)] <- 0

# Creating the Similarity matrix
similarity_matrix <- data.frame(matrix(nrow = random_samples_to_choose, ncol = 3))
colnames(similarity_matrix) <- c('MovieId', 'MovieName', 'CosineSimilarity')

similarity_matrix$MovieId <- random_movieIds
similarity_matrix$CosineSimilarity <- apply(movie_profile_matrix, 1, function(x) cosine(user_profile_vector, as.numeric(x)))
similarity_matrix$MovieName <- sapply(random_movieIds, function(x) as.character(movies[movies$movieId == x, 'title']))

# Ordering the Similarity matrix in decreasing order
similarity_matrix <- similarity_matrix[order(similarity_matrix$CosineSimilarity, decreasing = TRUE), ]

# Printing final required output
cat('User ID', user_profile_id, 'chose the following 10 movies:', toString(random_movieIds),'\n\n')
cat('Of these, the following 5 movies are recommended:\n\n')
print.AsIs(similarity_matrix[1:5, ], row.names = FALSE)

```

# Problem 2.3
```{r}
# Getting the movieIds watched by the user 191
user_profile_id_191 <- 191
test_movie_ids <- c(150, 296, 380, 590)

# Setting ratings of test movies for user191 to NA
actual_rating_by_user191 <- ratings[ratings$userId == user_profile_id_191, c('movieId', 'rating')]
movieIds_for_user191 <- actual_rating_by_user191$movieId
modified_ratings_of_user191 <- actual_rating_by_user191
modified_ratings_of_user191[is.element(modified_ratings_of_user191$movieId, test_movie_ids), 'rating'] <- NA

similar_users_to_user191 <- res[res$a == 'user191', ]
newrow <- c('user191', 'user568', '0.210526300000000')

top12_similar_users_to_user191 <- head(similar_users_to_user191[order(similar_users_to_user191$score, decreasing = TRUE), ], 13)
top12_similar_users_to_user191 <- top12_similar_users_to_user191[-c(4, 10), ]
top12_similar_users_to_user191 <- InsertRow(top12_similar_users_to_user191, newrow, 11)

```

# Problem 2.3 (a)
```{r}
# Getting the random similar users to user191
random_users_to_choose <- 5

random_users <- top12_similar_users_to_user191[sample(nrow(top12_similar_users_to_user191), size = random_users_to_choose, replace = FALSE), 'b']
random_users <- substring(random_users$b, 5)

# Making the empty Utility matrix
utility_matrix <- data.frame(matrix(nrow = random_users_to_choose+1, ncol = length(movieIds_for_user191)))
rownames(utility_matrix) <- c(random_users, as.character(user_profile_id_191))
colnames(utility_matrix) <- movieIds_for_user191

# Filling Utility matrix for random users
for (userId in random_users) {
  data_for_particular_user <- ratings[ratings$userId == userId, c('movieId', 'rating')]
  intersected_movieIds_for_particular_user <- intersect(movieIds_for_user191, data_for_particular_user$movieId)
  intersected_data_for_particular_user <- data_for_particular_user[is.element(data_for_particular_user$movieId, intersected_movieIds_for_particular_user), ]
  
  for (intersected_movieId in intersected_movieIds_for_particular_user) {
    utility_matrix[userId, as.character(intersected_movieId)] <- intersected_data_for_particular_user[intersected_data_for_particular_user$movieId == intersected_movieId, 'rating']
  }
}

# Filling Utility matrix for user191
utility_matrix[as.character(user_profile_id_191), ] <- modified_ratings_of_user191$rating

# Top 3 users (Neighbourhood = 3) based on Jaccard similarity from the 5 randomly chosen users
five_random_similar_users_to_user191 <- top12_similar_users_to_user191[is.element(top12_similar_users_to_user191$b, paste('user',random_users, sep = "")), ]
five_random_similar_users_to_user191 <- five_random_similar_users_to_user191[order(five_random_similar_users_to_user191$score, decreasing = TRUE), ]
top3_random_similar_uers_to_user191 <- five_random_similar_users_to_user191[1:3, ]

cat("USER ID ", user_profile_id_191, ", 5 random user IDs: ", toString(random_users), '\n\n', sep = "")
cat("Using user-user similarity, User ID", user_profile_id_191, "will rate the movies as follows:\n\n")

# Predicting ratings of test movies by user191 based on user-user similarity
error <- 0
for (test_movie_id in test_movie_ids) {
  numerator <- 0
  denominator <- 0
  
  for (userId in top3_random_similar_uers_to_user191$b) {
    
    substr_userId <- substring(userId, 5)
    rating <- utility_matrix[as.character(substr_userId), as.character(test_movie_id)]
    similarity <- as.numeric(top3_random_similar_uers_to_user191[top3_random_similar_uers_to_user191$b == userId, 'score'])
      
    if(!is.na(rating)) {
      numerator <- numerator + (similarity * rating)
    }
    denominator <- denominator + similarity
  }
  
  rating_by_user191 <- as.double(numerator/denominator)
  actual_rating <- as.double(actual_rating_by_user191[actual_rating_by_user191$movieId == test_movie_id, 'rating'])
  error <- error + ((rating_by_user191 - actual_rating)^2)
  cat(test_movie_id, ": ", rating_by_user191, '\n', sep = "")
}

cat("RMSE: ", sqrt(error/length(test_movie_ids)), sep = "")

```

# Problem 2.3 (b)
```{r}
# Getting the random similar users
random_users_2 <- top12_similar_users_to_user191[sample(nrow(top12_similar_users_to_user191), size = random_users_to_choose, replace = FALSE), 'b']
random_users_2 <- substring(random_users_2$b, 5)

# Making the empty Utility matrix
utility_matrix_2 <- data.frame(matrix(nrow = length(movieIds_for_user191), ncol = random_users_to_choose+1))
rownames(utility_matrix_2) <- movieIds_for_user191
colnames(utility_matrix_2) <- c(random_users_2, as.character(user_profile_id_191))

# Filling Utility matrix for random users
for (userId in random_users_2) {
  data_for_particular_user <- ratings[ratings$userId == userId, c('movieId', 'rating')]
  intersected_movieIds_for_particular_user <- intersect(movieIds_for_user191, data_for_particular_user$movieId)
  intersected_data_for_particular_user <- data_for_particular_user[is.element(data_for_particular_user$movieId, intersected_movieIds_for_particular_user), ]
  
  for (intersected_movieId in intersected_movieIds_for_particular_user) {
    utility_matrix_2[as.character(intersected_movieId), userId] <- intersected_data_for_particular_user[intersected_data_for_particular_user$movieId == intersected_movieId, 'rating']
  }
}

# Filling Utility matrix for user191
utility_matrix_2[, as.character(user_profile_id_191)] <- modified_ratings_of_user191$rating

cat("USER ID ", user_profile_id_191, ", 5 random user IDs: ", toString(random_users_2), '\n\n', sep = "")
cat("Using item-item similarity, User ID", user_profile_id_191, "will rate the movies as follows:\n\n")

# Matrix with mean substracted from all rows
mean_substracted_matrix <- sweep(utility_matrix_2, 1, apply(utility_matrix_2, 1, function(x) mean(x, na.rm = TRUE)))
mean_substracted_matrix[is.na(mean_substracted_matrix)] <- 0

# Predicting ratings of test movies by user191 based on item-item similarity
error <- 0
for (test_movie_id in test_movie_ids) {
  numerator <- 0
  denominator <- 0
  
  # Row with mean subtracted values for test movie
  test_row <- as.double(mean_substracted_matrix[as.character(test_movie_id), ])
  
  # Finding cosine similarity between test movie and all other movie rows
  mean_cosine_similarity <- as.data.frame(apply(mean_substracted_matrix, 1, function(x) cosine(test_row, as.double(x))))
  colnames(mean_cosine_similarity) <- c('CosineSimilarity')
  mean_cosine_similarity$movieId <- movieIds_for_user191
  mean_cosine_similarity <- mean_cosine_similarity[order(mean_cosine_similarity$CosineSimilarity, decreasing = TRUE), ]
  
  # Selecting top 3 similarities
  top3_mean_cosine_similarity <- mean_cosine_similarity[2:4, ]
  
  # Calculating rating by user191
  for (movieId in top3_mean_cosine_similarity$movieId) {
    
    rating <- utility_matrix_2[as.character(movieId), as.character(user_profile_id_191)]
    similarity <- top3_mean_cosine_similarity[top3_mean_cosine_similarity$movieId == movieId, 'CosineSimilarity']
      
    if(!is.na(rating)) {
      numerator <- numerator + (similarity * rating)
    }
    denominator <- denominator + similarity
  }
  
  rating_by_user191 <- as.double(numerator/denominator)
  actual_rating <- as.double(actual_rating_by_user191[actual_rating_by_user191$movieId == test_movie_id, 'rating'])
  error <- error + ((rating_by_user191 - actual_rating)^2)
  cat(test_movie_id, ": ", rating_by_user191, '\n', sep = "")
}

cat("RMSE: ", sqrt(error/length(test_movie_ids)), sep = "")

```